### 2020-09-11 rescience/metadata.tex

% DO NOT EDIT - automatically generated from metadata.yaml

\def \codeURL{https://github.com/cboettig/fluctationDomains}
\def \codeDOI{}
\def \codeSWH{}
\def \dataURL{}
\def \dataDOI{}
\def \editorNAME{}
\def \editorORCID{}
\def \reviewerINAME{}
\def \reviewerIORCID{}
\def \reviewerIINAME{}
\def \reviewerIIORCID{}
\def \dateRECEIVED{}
\def \dateACCEPTED{}
\def \datePUBLISHED{}
\def \articleTITLE{[Re] Fluctation Domains}
\def \articleTYPE{Reproduction}
\def \articleDOMAIN{Ten Years Reproducibility Challenge}
\def \articleBIBLIOGRAPHY{bibliography.bib}
\def \articleYEAR{2020}
\def \reviewURL{}
\def \articleABSTRACT{A stochastic ensemble simulation of evolutionary dynamics has been successfully reproduced. Simulation code was in C and could be called from wrappers implemented as an R package.  Reproduction required minor modification to the configuration files for compatibility with newer versions of the GNU Scientific Library.}
\def \replicationCITE{Carl Boettiger, Jonathan Dushoff, Joshua S Weitz (2010). Fluctuation domains in adaptive evolution, Theoretical Population Biology 77 (1) 6-13. doi:10.1016/j.tpb.2009.10.003.}
\def \replicationBIB{Boettiger2010}
\def \replicationURL{https://arxiv.org/abs/1004.4233}
\def \replicationDOI{https://doi.org/10.1016/j.tpb.2009.10.003}
\def \contactNAME{Carl Boettiger}
\def \contactEMAIL{cboettig@berkeley.edu}
\def \articleKEYWORDS{ten years reproducibility challenge, R, C}
\def \journalNAME{ReScience C}
\def \journalVOLUME{5}
\def \journalISSUE{2}
\def \articleNUMBER{}
\def \articleDOI{}
\def \authorsFULL{Carl Boettiger}
\def \authorsABBRV{C. Boettiger}
\def \authorsSHORT{Boettiger}
\title{\articleTITLE}
\date{}
\author[1,\orcid{https://orcid.org/0000-0002-1642-628X}]{Carl Boettiger}
\affil[1]{University of California, Berkeley, 130 Mulford Hall, Berkeley, CA 94720-3114 USA}